---
title: "9041 Assignment 2"
author: Ali Naeij 
date: 10/02/2026
format: 
    html:
      toc: true
      code-fold: true
      code-summary: "Show code"
      df-print: paged
execute:
    echo: false
    warning: false
---


```{r, load libraries}
#| output: FALSE


library(tidyverse)
library(readxl)
library(psych)
library(gt)
library(janitor)
```
```{r}
#| output: false

#loading and inspecting data
data <- read.csv("Copy of Attachment_Anxiety_Data.csv")

is.numeric(data$Age.group)

mutated.data <- data |>
  mutate(Gender.Char = case_when(Gender == 1 ~ "Male", 
                                 Gender == 2 ~ "Female")) |>
  mutate(Relationship.Char = case_when(Relationship == 1 ~ "Married",
                                       Relationship %in% c(2,5)  ~ "In a Relationship",
                                       Relationship == 4 ~ "Divorced",
                                       Relationship == 6 ~ "Never Been in a Relationship",
                                       Relationship == 7 ~ "Prefer Not to Say")) |>
  mutate(Ethnicity.Char = case_when (Ethnicity == 1 ~ "Caucasian",
                                     Ethnicity == 2 ~ "Asian",
                                     Ethnicity == 5 ~ "Other",
                                     Ethnicity == 6 ~ "Prefer Not to Say")) |>
  mutate(Age.Char = case_when (Age.group == 1 ~ "18-24",
                               Age.group == 2 ~ "24-34",
                               Age.group == 3 ~ "35-44",
                               Age.group == 4 ~ "45-54",
                               Age.group == 5 ~ "55-64"))
  
```


```{r, create keys}
#| output: FALSE

SA.Keys <- list(Social.Anxiety=c("SA_1","SA_2","SA_3","SA_4","-SA_5","SA_6","SA_7","SA_8",
                             "-SA_9","SA_10","-SA_11","SA_12","SA_13","SA_14","SA_15", 
                             "SA_16","SA_17","SA_18","SA_19"))
AA.Keys <- list(Attachment.Anxiety=c("AA_1", "AA_2", "AA_3", "AA_4", 
                                  "AA_5", "AA_6", "AA_7", "AA_8", "AA_9"))
SEst.Keys <- list(Rosenberg.SelfEsteem = c("SEst_1", "-SEst_2", "SEst_3", "SEst_4","-SEst_5",
                                           "-SEst_6", "SEst_7", "-SEst_8", "-SEst_9", "SEst_10"))

```

```{r, score items}
#| output: FALSE

Scores.SA <- scoreItems(SA.Keys, mutated.data, min = 0, max = 5)
Scores.AA <- scoreItems(AA.Keys, mutated.data, min =1, max = 7)
scores.SEst <- scoreItems(SEst.Keys, mutated.data, min = 1, max = 4)
```

```{r, combine scores}
#| output: FALSE

Scores.SA.df <- as.data.frame(Scores.SA$scores)
Scores.AA.df <- as.data.frame(Scores.AA$scores)
scores.SEst.df <- as.data.frame(scores.SEst$scores)

allscores.df <- cbind(Scores.SA.df, Scores.AA.df, scores.SEst.df)
```

```{r, combine dataframes}
#| output: FALSE

final.data <- cbind(mutated.data, allscores.df)
```

```{r, simplyfying the datafram}
#| output: false


simple.final.data <- final.data |> 
  select(Gender.Char, Age.Char, Ethnicity.Char, Relationship.Char, Social.Anxiety, Attachment.Anxiety, Rosenberg.SelfEsteem )
```


```{r, creating a demographic table}
#| output: false


demographic.table <- simple.final.data |>
  tabyl(Gender.Char, Relationship.Char)

demographic.gt <- demographic.table |> 
  gt() |> 
  tab_header(title = "Participant Relationship Status by Age")
```

```{r, score table}
#| output: false


#step 1: summary statistics for the 3 scales

summary <- simple.final.data |> 
  select(Social.Anxiety, Attachment.Anxiety, Rosenberg.SelfEsteem) |>
  describe() |> 
  as.data.frame()

#step 2: sample size
sample.size <- nrow(simple.final.data)

#step 3: scales column
summary$Scales <- c("Social Interaction Anxiety", "Attachment Anxiety", "Rosenberg Self-Esteem")

#step : alphas, rounded to 2 decimals
alphas <- round(c(Scores.SA$alpha,
                  Scores.AA$alpha,
                  scores.SEst$alpha), 2)

#simplifying the table 

summary.table <- summary |> 
  select(Scales, mean, median, sd, range)

summary.table <- cbind(summary.table, Alpha = alphas)

summary.gt <- summary.table |>
  gt() |>
  tab_header("Participant Descriptive Stats")

```

```{r, filtering by variable}
#| output: false


female.data <- simple.final.data |> 
  filter(Gender.Char == "Female")

nrow(female.data)
```
```{r, filtered data and scored table}
#| output: false

#summary stats for the filtered group
summary.female <- female.data |> 
  select(Social.Anxiety, Attachment.Anxiety, Rosenberg.SelfEsteem) |>
  describe() |>
  as.data.frame()

sample.size.female <- nrow(female.data)

summary.female$Scales <- c("Social Interaction Anxiety",
                           "Attachment Anxiety",
                           "Rosenberg Self-Esteem")

#Alphas for the female only dataset

female.items <- final.data |> filter(Gender.Char == "Female")

Scores.SA.female <- scoreItems(SA.Keys, female.items, min = 0, max= 5)
Scores.AA.female <- scoreItems(AA.Keys, female.items, min = 1, max = 7)
Scores.SEst.female <- scoreItems(SEst.Keys, female.items, min = 1, max = 4)

alphas.female <- round(c(Scores.SA.female$alpha,
                         Scores.AA.female$alpha,
                         Scores.SEst.female$alpha), 2)

summary.table.female <- summary.female |> 
  select(Scales, mean, median, sd, range)

summary.table.female <- cbind(summary.table.female, Alpha = alphas.female)

summary.female.gt <- summary.table.female |> 
  gt() |> 
  tab_header(title = "Female Participant Descriptive Statistics")

```

## Intro

The data in this document is a compilation of self-reported measures from 63 different individuals, assessing Social Anxiety, Attachment Anxiety, Self-Esteem, obtained through responses to a set of questions. 

Participants also reported on Age, Gender, Relationship Status, and Ethnicity

## Methods

### Participants

The sample consisted of `r sample.size` participants. Their demographic characteristics are shown in Table 1.

```{r show-demographics}
demographic.gt
```


### Measures

Attachment anxiety was measured using a shortened version of the anxiety subscale of the Experiences in Close Relationships scale, consisting of the nine items that loaded most strongly on the anxiety factor in Sibley, Fischer and Liu (2005). Items were rated from 1 to 7 (α = `r alphas[2]`). Social anxiety was measured with the 20-item Social Interaction Anxiety Scale (α = `r alphas[1]`), and self-esteem with the 10-item Rosenberg Self-Esteem Scale, rated from 1 to 4 (α = `r alphas[3]`).

### Data Analysis Strategy

Before scoring, negatively worded items were reverse-keyed so that higher scores reflected more of each construct. This applied to items 5, 9 and 11 of the Social Anxiety Scale and items 2, 5, 6, 8 and 9 of the Rosenberg Self-Esteem Scale; no attachment anxiety items were reversed. Scale scores were calculated as the mean of each participant's item responses using `scoreItems()` from the psych package, and internal consistency was assessed with Cronbach's alpha.


## Results

Descriptive statistics for the full sample are shown in Table 2.

```{r show-summary}
summary.gt
```

Participants reported a mean social interaction anxiety score of `r round(summary.table$mean[1], 2)`, a mean attachment anxiety score of `r round(summary.table$mean[2], 2)`, and a mean self-esteem score of `r round(summary.table$mean[3], 2)`.

Table 3 shows the same statistics for female participants only (n = `r sample.size.female`).

```{r show-summary-female}
summary.female.gt
```

Among female participants, mean scores were `r round(summary.table.female$mean[1], 2)` for social interaction anxiety, `r round(summary.table.female$mean[2], 2)` for attachment anxiety, and `r round(summary.table.female$mean[3], 2)` for self-esteem.



























